---
title: "Risk_Analysis"
author: "Ojo Ilesanmi and Somto Mommah"
date: "6/9/2020"
output: html_document
---

#Libraries
```{r}
library('tidyverse')
library('DMwR')
library('randomForest')
library('rpart')
library('cluster')
library('recipes')
library('gridExtra')
library('caret')
library('xgboost')
```

```{r}
credit_data <- read.csv('C:/Users/USER/Documents/Risk Analysis/train_loan.csv')
```

```{r}
str(credit_data)
```

```{r}
sapply(credit_data, function(x) sum(is.na(x)))
```
#Replacing missing values in Loan_Amount_Term with the Mean
```{r}
credit_data$Loan_Amount_Term<- floor(ifelse(is.na(credit_data$Loan_Amount_Term), mean(credit_data$Loan_Amount_Term, na.rm=TRUE), credit_data$Loan_Amount_Term))
```

#Replacing missing values in credit history with mode
```{r}
Mode = function(x){
    ta = table(x)
    tam = max(ta)
    if (all(ta == tam))
         mod = NA
    else
         if(is.numeric(x))
    mod = as.numeric(names(ta)[ta == tam])
    else
         mod = names(ta)[ta == tam]
    return(mod)
}

Credit_History_mode <- Mode(credit_data$Credit_History)


credit_data$Credit_History <- ifelse(is.na(credit_data$Credit_History),Credit_History_mode,credit_data$Credit_History)

```
#Replacing missing values in Dependents with mean
```{r}
credit_data$Dependents<- floor(ifelse(is.na(credit_data$Dependents), mean(credit_data$Dependents, na.rm=TRUE), credit_data$Dependents))
```

#Replacing missing values in Loan Amount with mean
```{r}
credit_data$LoanAmount<- floor(ifelse(is.na(credit_data$LoanAmount), mean(credit_data$LoanAmount, na.rm=TRUE), credit_data$LoanAmount))
```

#Converting Dollars to Naira in LoanAmount
```{r}
credit_data$LoanAmount <- credit_data$LoanAmount * 410
```

#converting LoanAmount from Yearly basis to Monthly basis
```{r}
credit_data$LoanAmount <- floor((credit_data$LoanAmount) / 12)
```
#Dropping column #Loan_ID it won't be needed in our analysis and prediction
```{r}
credit_data$Loan_ID <- NULL
```

```{r}
p1 <- ggplot(credit_data, aes(x=Education)) + ggtitle("Educational Qualifications") + xlab("Education") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p2 <- ggplot(credit_data, aes(x=Married)) + ggtitle("Marital Status") + xlab("Marital Status") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p3 <- ggplot(credit_data, aes(x=Self_Employed)) + ggtitle("Employment Status ") + xlab("Self Employed or Not") + 
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p4 <- ggplot(credit_data, aes(x=Gender)) + ggtitle("Gender") + xlab("Male") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()

p5 <- ggplot(credit_data, aes(x=Dependents)) + ggtitle("Dependants") + xlab("Number of Dependants") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p6 <- ggplot(credit_data, aes(x=Credit_History)) + ggtitle("Credit History") + xlab("Credit History") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
p7 <- ggplot(credit_data, aes(x=Property_Area)) + ggtitle("Location of Property") + xlab("Property Location") +
  geom_bar(aes(y = 100*(..count..)/sum(..count..)), width = 0.5) + ylab("Percentage") + coord_flip() + theme_minimal()
grid.arrange(p1, p2, p3, p4, p5, p6, p7, ncol=3)

```

```{r}
creditdata_noimp<-knnImputation(credit_data, k = 5, scale = T, meth = "weighAvg",
distData = NULL)
nrow(creditdata_noimp)
```

```{r}
p8 <- ggplot(credit_data, aes(x= ApplicantIncome)) + ggtitle("Applicant's Income") + xlab("Applicant Income") + 
  geom_histogram() + theme_minimal()
p9 <- ggplot(credit_data, aes(x=LoanAmount)) + ggtitle("Loan Amount") + xlab("Loan Amount") + 
  geom_histogram() + theme_minimal()
p10 <- ggplot(credit_data, aes(x=Loan_Amount_Term)) + ggtitle("Loan Amount Term") + xlab("Loan Amount Term") + 
  geom_histogram() + theme_minimal()
p11 <- ggplot(credit_data, aes(x=CoapplicantIncome)) + ggtitle("CoApplicant's Income") + xlab("CoApplicant's Income") + 
  geom_histogram() + theme_minimal()
grid.arrange(p8,p9, p10, p11)

```
#Splitting data for modelling (train and test)
```{r}
split<-sample(nrow(transform_3), round(nrow(transform_3)*0.8))
trainingdata <-transform_3[split,]
testdata <- transform_3[-split,]
dim(trainingdata)
dim(testdata)
```

```{r}
dummy <- dummyVars('~.', data = credit_data, sep = '.')
transform <- data.frame(predict(dummy, newdata = credit_data))
view(transform)

transform_2 <- transform[,-c(1,2,4,5,9,10,11,18,21)]
view(transform_2)

transform_3 <- transform_2 %>% rename(
  Gender = Gender.Male,
  Married = Married.Yes,
  Education = Education.Graduate,
  Self_Employed = Self_Employed.Yes,
  Loan_Status = Loan_Status.Y
    )
    
View(transform_3)
```

```{r}
transform_3$Loan_Status <- as.factor(transform_3$Loan_Status)
transform_3$Dependents <- as.factor(transform_3$Dependents)
transform_3$Married <- as.factor(transform_3$Married)
transform_3$Education <- as.factor(transform_3$Education)
transform_3$Gender <- as.factor(transform_3$Gender)
transform_3$Self_Employed <- as.factor(transform_3$Self_Employed)
transform_3$Credit_History <- as.factor(transform_3$Credit_History)
```

```{r}
cross_validation <- trainControl(method = "repeatedcv", repeats = 1,number = 3)
```

```{r}
logistic_regression<- train(form=Loan_Status~., data=trainingdata,method="glm", family="binomial",trControl=cross_validation)
```

```{r}
logistic_regression_prediction<-predict(logistic_regression,transform_5,type="raw")
logistic_regression_confusion_matrix<-confusionMatrix(logistic_regression_prediction, transform_5$Loan_Status)
```

```{r}
#result <- confusionMatrix(prediction, truth)
precision <- logistic_regression_confusion_matrix$byClass['Pos Pred Value']    
recall <- logistic_regression_confusion_matrix$byClass['Sensitivity']
```

```{r}
svm<- train(form=Loan_Status~., data=trainingdata, method="svmLinear", trControl=cross_validation, na.action=na.roughfix )
```

```{r}
predsvm<-predict(svm,testdata,type="raw")
confusionMatrix(predsvm, testdata$Loan_Status)
```

```{r}
mtry <- sqrt(ncol(trainingdata))
tunegrid <- expand.grid(.mtry=mtry)
rf<- train(form=Loan_Status~., data=trainingdata,method="rf", metric="Accuracy",tuneGrid=tunegrid, trControl=cross_validation)
```

```{r}
predrf<-predict(rf,testdata,type="raw")
confusionMatrix(predrf, testdata$Loan_Status)
```

```{r}
xgb.grid <- expand.grid(nrounds = 100,eta=0.3, gamma=0, max_depth=3, min_child_weight=1, subsample=1, colsample_bytree=1)
```

```{r}
xgb_model <-train(Loan_Status ~ .,data = trainingdata, method="xgbTree",trControl=cross_validation, tuneGrid=xgb.grid,nthread =4)
```

```{r}
predxgb<-predict(xgb_model,testdata,type="raw")
confusionMatrix(predxgb, testdata$Loan_Status)
```

```{r}
test_loan <- read.csv('C:/Users/USER/Documents/Risk Analysis/train_loan.csv')
```

```{r}
test_loan$Loan_Amount_Term<- floor(ifelse(is.na(test_loan$Loan_Amount_Term), mean(test_loan$Loan_Amount_Term, na.rm=TRUE), test_loan$Loan_Amount_Term))
```

```{r}

Mode = function(x){
    ta = table(x)
    tam = max(ta)
    if (all(ta == tam))
         mod = NA
    else
         if(is.numeric(x))
    mod = as.numeric(names(ta)[ta == tam])
    else
         mod = names(ta)[ta == tam]
    return(mod)
}

Credit_History_mode <- Mode(test_loan$Credit_History)


test_loan$Credit_History <- ifelse(is.na(test_loan$Credit_History),Credit_History_mode,test_loan$Credit_History)
```

```{r}
test_loan$Dependents<- floor(ifelse(is.na(test_loan$Dependents), mean(test_loan$Dependents, na.rm=TRUE), test_loan$Dependents))
```

```{r}
test_loan$LoanAmount<- floor(ifelse(is.na(test_loan$LoanAmount), mean(test_loan$LoanAmount, na.rm=TRUE), test_loan$LoanAmount))
```

```{r}
test_loan$LoanAmount <- test_loan$LoanAmount * 410
```

```{r}
test_loan$LoanAmount <- floor((test_loan$LoanAmount) / 12)
```

```{r}
test_loan$Loan_ID <- NULL
```

```{r}
dummy <- dummyVars('~.', data = test_loan, sep = '.')
transform_1 <- data.frame(predict(dummy, newdata = test_loan))
#view(transform)

transform_4 <- transform_1[,-c(1,2,4,5,9,10,11,18,21)]
#view(transform_2)

transform_5 <- transform_4 %>% rename(
  Gender = Gender.Male,
  Married = Married.Yes,
  Education = Education.Graduate,
  Self_Employed = Self_Employed.Yes,
  Loan_Status = Loan_Status.Y
    )
    
#View(transform_3)
```

```{r}
transform_5$Loan_Status <- as.factor(transform_5$Loan_Status)
transform_5$Dependents <- as.factor(transform_5$Dependents)
transform_5$Married <- as.factor(transform_5$Married)
transform_5$Education <- as.factor(transform_5$Education)
transform_5$Gender <- as.factor(transform_5$Gender)
transform_5$Self_Employed <- as.factor(transform_5$Self_Employed)
transform_5$Credit_History <- as.factor(transform_5$Credit_History)
```